---
title: "Project 4"
author: "Lauren Forti"
date: "11/7/2022"
output: html_document
---



### Read in dataset
```{r}
library(readr)

# read in data
Batteries_Life <- read.csv("Batteries_Life.csv")

# display first 6 rows
head(Batteries_Life, 6)

# convert brand and device to factors
Batteries_Life$Brand <- as.factor(Batteries_Life$Brand)
Batteries_Life$Device <- as.factor(Batteries_Life$Device)
```



### ANOVA Test: Main Effects/Interaction
```{r}
# perform ANOVA
Batteries_Life.factors.aov <- aov(Life~Brand*Device, Batteries_Life)

# output results
summary(Batteries_Life.factors.aov)

# get F critical val for Brand
qf(c(0.05),1,8,lower.tail=FALSE)

# get F critical val for Device
qf(c(0.05),2,8,lower.tail=FALSE)
```

The main effects for Brand and Device are significant.  

At α=0.05, the critical value of F for Brand is 5.32. The F Value for Brand is 9.33. Since the critical value of F = 5.32 is less than F = 9.33, the hypothesis should not be rejected. Brand does affect the battery life. 

At α=0.05, the critical value of F for Device is 4.46. The F Value for Device is 130.75. Since the critical value of F = 4.46 is less than F = 130.75, the hypothesis should not be rejected. Device does affect the battery life. 



### Interaction Plot
```{r}
# make interaction plot
with(Batteries_Life, interaction.plot(Brand, Device, Life, type="b", col=10:12), pch=19, fixed=T, xlab="Brand", ylab="Life")
```


### ANOVA Test: Interaction
```{r}
# output results
summary(Batteries_Life.factors.aov)

# get F critical val for interaction
qf(c(0.05),2,6,lower.tail=FALSE)
```

The interaction between Brand and Device is not significant.  

At α=0.05, the critical value of F for interaction between Brand and Device is 5.14. The F Value for the interaction is 0.476. Since the critical value of F = 5.14. is greater than F = 0.476, the hypothesis should be rejected. There is little interaction between Brand and Device. Brand and Device separately have statistical significance.



### Residual Plots
```{r}
library(ggplot2)
library(lmtest)

# plot normal qq plot
plot(Batteries_Life.factors.aov, which=2)

# plot scale-location plot
plot(Batteries_Life.factors.aov, which=3)
# breusch-pagan test
bptest(Batteries_Life.factors.aov)

# plot residuals vs fitted plot
plot(Batteries_Life.factors.aov, which=1)

# append residuals to df
Batteries_Life$residual <- Batteries_Life.factors.aov$residuals

# plot residuals vs brand
ggplot(Batteries_Life, aes(x=Brand, y=residual)) +
  geom_point(size=2, shape=23)+
  labs(title = "Plot of residuals vs. Brand")

# plot residuals vs device
ggplot(Batteries_Life, aes(x=Device, y=residual)) +
  geom_point(size=2, shape=23)+
  labs(title = "Plot of residuals vs. Device")
```

From these plots, it is safe to assume that the model is adequate.  

**Normal QQ**  
The residuals loosely form a line that deviates at the beginning and end. This indicates that there are extreme values at the tails. It is safe to assume that the distribution is relatively normal.  


**Scale-Location**  
The line is slightly horizontal. A Breusch-Pagan Test results in a p-value of 0.035, which is smaller than 0.05. This indicates that it is safe to assume homoscedasticity.  


**Residuals vs Fitted**  
The line is horizontal and the residuals are evenly spread around the horizon line. This indicates that there is a linear relationship present.  



### Recommendations
```{r}
library(emmeans)

# get mean comparisons
emmeans(Batteries_Life.factors.aov, pairwise ~ Brand | Device)
```

Brand B should be recommended over Brand A.  

Brand B tends to have a longer battery life across devices looking at the interaction plot and the mean comparisons.  
\
\
\