---
title: 'Project 5: General Factor Design'
author: "Lauren Forti"
date: "11/15/2022"
output: html_document
---


### Read in dataset
```{r warning=FALSE, message=FALSE}
library(readr)
library(dplyr)

# read in data
Yield <- read.csv("Yield.csv")

# change Day3 to Day2
Yield$Blocks[Yield$Blocks=="Day3"] <- "Day2"

# display first 6 rows
head(Yield, 6)

# convert to factors
Yield$Blocks <- as.factor(Yield$Blocks)
Yield$Temperature <- as.factor(Yield$Temperature)
Yield$Pressure <- as.factor(Yield$Pressure)
```



### Explore Data Graphically
```{r warning=FALSE, message=FALSE}
library(ggplot2)
library(gridExtra)
library(grid)

# get averages for stripcharts
treatment_p.means <- with(Yield, tapply(Yield,Pressure,mean))
treatment_t.means <- with(Yield, tapply(Yield,Temperature,mean))
treatment_b.means <- with(Yield, tapply(Yield,Blocks,mean))

# pressure stripchart
stripchart(Yield~Pressure, data=Yield, vertical=TRUE, pch=1, main="Yield vs Pressure", xlab="Pressure", ylab="Yield")
stripchart(as.numeric(treatment_p.means)~as.numeric(names(treatment_p.means)),pch="x",cex=1.5,vert=T,add=T)

# temp stripchart
stripchart(Yield~Temperature, data=Yield, vertical=TRUE, pch=1, main="Yield vs Temperature", xlab="Temperature", ylab="Yield")
stripchart(as.numeric(treatment_t.means)~(names(treatment_t.means)),pch="x",cex=1.5,vert=T,add=T)

# blocks stripchart
stripchart(Yield~Blocks, data=Yield, vertical=TRUE, pch=1, main="Yield vs Blocks", xlab="Temperature", ylab="Blocks")
stripchart(as.numeric(treatment_b.means)~(names(treatment_b.means)),pch="x",cex=1.5,vert=T,add=T)

# plot boxplots
boxplot(Yield~Blocks, data=Yield)
boxplot(Yield~Temperature, data=Yield)
boxplot(Yield~Pressure, data=Yield)
```

From the stripcharts and boxplots, it seems like all three factors (Blocks, Temperature, and Pressure) effect the yield.   



### Are Blocking Factors Necessary?
```{r}
# perform ANOVA
Yield.full.aov <- aov(Yield~Blocks+Temperature+Pressure+Temperature*Pressure, Yield)

# output results
summary(Yield.full.aov)
```

Blocking factors are necessary for this experiment. The p-value is 0.0011, which is lower than α=0.05. This means that blocking factors are significant.  



### ANOVA (Main Effects)
```{r}
# get means
Temperature_mean <- Yield %>% group_by(Temperature) %>% summarise(Yield_mean=mean(Yield))
Pressure_mean <- Yield %>% group_by(Pressure) %>% summarise(Yield_mean=mean(Yield))
Blocks_mean <- Yield %>% group_by(Blocks) %>% summarise(Yield_mean=mean(Yield))

# plot standard error of means
ggplot1 <- ggplot(Temperature_mean, aes(x=Temperature, y=Yield_mean, group=1)) + 
     geom_line(color="cadetblue3", size=1) +
     geom_point(color="cadetblue4", size=1.2)
ggplot2 <-ggplot(Pressure_mean, aes(x=Pressure, y=Yield_mean, group=1)) + 
    geom_line(color="brown3", size=1) +
    geom_point(color="brown4", size=1.2)
ggplot3 <-ggplot(Blocks_mean, aes(x=Blocks, y=Yield_mean, group=1)) + 
    geom_line(color="chartreuse3", size=1) +
    geom_point(color="chartreuse4", size=1.2)

# display plots
grid.arrange(ggplot1,ggplot2,ggplot3, nrow = 1, top=textGrob("Main Effect Plots",gp=gpar(fontsize=18,font=3)))

# perform ANOVA
Yield.main.aov <- aov(Yield~Temperature*Pressure+Blocks, Yield)

# output ANOVA results
summary(Yield.main.aov)

# get F critical val for Temperature and Pressure
qf(c(0.05),2,8,lower.tail=FALSE)
```

The main effects for Temperature and Pressure are significant.  

At α=0.05, the critical value of F for Temperature is 4.459. The F Value for Temperature is 93.981. Since the critical value of F = 4.459 is less than F = 93.981, the hypothesis should not be rejected. Temperature does affect the yield.  

At α=0.05, the critical value of F for Pressure is 4.459. The F Value for Temperature is 5.184. Since the critical value of F = 4.459 is less than F = 5.184, the hypothesis should not be rejected. Pressure does affect the yield.  



### Interaction Plot
```{r}
# make interaction plot
with(Yield, interaction.plot(Temperature,Pressure,Yield,type="b",col=10:12
                                    ,pch=19, fixed=T,xlab="Temperature",ylab="Average Yield"))
title (main = "Interaction Plot")
```



### ANOVA (Main Effects and Interactions)
```{r}
# output ANOVA results
summary(Yield.full.aov)

# get F critical val for Temperature and Pressure interaction
qf(c(0.05),4,8,lower.tail=FALSE)
```

The interaction between Temperature and Pressure is not significant.  

At α=0.05, the critical value of F for the interaction between Temperature and Pressure is 3.838. The F Value for the interaction is 2.095. Since the critical value of F = 3.838 is greater than F = 2.095, the hypothesis should be rejected. The interaction between Temperature and Pressure is not significant.  
\
\
\