---
title: "Project 1 - Photoresist Thickness Study"
author: "Lauren Forti"
date: "10/18/2022"
output: html_document
---


# Load in the Data
```{r}
Temp95 <- c(11.176, 7.089, 8.097, 11.739, 11.291, 10.759, 6.467, 8.315)
Temp100 <- c(5.263, 6.748, 7.461, 7.015, 8.133, 7.418, 3.772, 8.963)

# make df
df <- data.frame(Temp95, Temp100)

# output df
df
```

# T-test
```{r}
# conduct t-test
t.test(Temp95, Temp100, var.equal=TRUE)
```



# Problem 1
**Is there evidence to support the claim that the higher baking temperature results in wafers with a lower mean photoresist thickness? Use α = 0.05.**

Yes, there is evidence to support the claim that the higher baking temperature results in wafers with a lower mean photoresist thickness. The null hypothesis should not be rejected at significance level α = 0.05. This is because the p-value is 0.01812 which is lower than the significance level.



# Problem 2
**What is the P-value for the test conducted?**

The p-value is 0.01812.



# Problem 3
**Find a 95% confidence interval on the difference in means. Provide a practical interpretation of this interval.**

The 95% confidence interval is [0.4995743, 4.5404257]. This means that the maximum difference between the means can be as low as 0.4995743 or as high as 4.5404257.



# Problem 4
**Draw dot diagrams and Box-and-whisker plot to assist in interpreting the results from this experiment.**

```{r}
# plot dot diagram
df.means <- as.numeric(apply(df,2,mean))
stripchart(df, xlab=expression("Thickness (kA)"),pch=19)
arrows(df.means,rep(1.5,2),df.means,c(1.1,1.9),length=.1)

# plot boxplot
boxplot(df, ylab="Thickness (kA)", las=1)
```
